### 2005-11-21 Documentation/part1.tex
*** empty log message ***

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Simulating non-standard physics}
\label{chapt:nsphysics}
\index{norm}{Non-standard physics}

In this chapter, we discuss how to simulate non-standard physics with \GLOBES , \ie, physics
beyond three-flavor neutrino oscillations. For the first time, this feature was used in \Ref~\cite{Blennow:2005yk},
and it has become a standard feature of \GLOBES\ starting from version 3.0 (CHECK). Since the 
computation of oscilltion probabilities, and therefore the flavor transition probabilities, is the core basic
element of \GLOBES , the introduction of non-standard physics requires that one produce a new version branch
of the software, \ie, the software has to be re-compiled after the modification of the probability engine.
Below, we will demonstrate how and where to do the necessary mofifications, and how to simulate non-standard
physics in the application software using the ``new'' \GLOBES .

\section{Modification of \GLOBES }

In order to modify \GLOBES , first install a current version of GLOBES (version larger than X.X.X CHECK) as
a new branch, \ie, in a new directory, and rename the directory \GLOBES\ expanded into. 
Using multiple versions of \GLOBES , it is, in general, advisable
to specify the installation directory of all used versions manually (see ``Installation without
root privilege'' in {\tt INSTALL} file). For instance, if the current ``standard'' version of \GLOBES\ is
installed in {\tt GLOBES/Versions}, you may use {\tt GLOBES/NSVersion} for the non-standard version
specified in {\tt configure}. Do not forget to rename the directory \GLOBES\ tar-ball expanded into: For instance,
change the name from {\tt globes-3.0.0} to {\tt globes-3.0.0nsi}. 
Then you can modify the source code in the sub-directories of {\tt globes-3.0.0nsi}, such as {\tt globes-3.0.0nsi/source}.
In particular, we will be interested in modifying {\tt glb\_probability.cc}, which contains the
Hamiltonian in the routine {\tt ComputeMixingMatrix}.
%
In general, the steps to introduce non-standard physics as follows:
\begin{enumerate}
\item
 Change the number of oscillation parameters in {\tt GLB\_OSCP} from six (standard) to the
 new number $6+N$, where $N$ is the number of non-standard parameters. 
 This change has to be done in the header file {\tt globes.h} in the sub-directory {\tt globes}
 where the tar-ball was extracted into,  {\em as well as} in the {\tt include/globes} sub-directory
 where the non-standard version was installed into (such as {\tt GLOBES/NSVersion/include/globes}).
\item 
 Modify {\tt glb\_probability.cc} to include the non-standard effects. The parameters
 can be accessed by {\tt nsparams[0]}, {\tt nsparams[1]}, ..., {\tt nsparams[N-1]}.
\item
 Re-compile the software using \\
 {\tt make} \\
 {\tt make install}
\end{enumerate}
Now your version of \GLOBES\ uses $6+N$ transition parameters, the six standard oscillation parameters,
and $N$ additional ones.

\section{Using a non-standard version of \GLOBES }

Using more than six transition parameters in \GLOBES , you will have to maintain the additional parameters. 
For example, there are no standard routines in \GLOBES\ which define projections including more than six parameters, 
which means that you should only use {\tt glbChiNP} further on. In addition, you can still use {\tt glbDefineParams}, but this 
will only access the six standard parameters. You will need to set the additional ones manually using {\tt glbSetOscParams}. 
Do not forget to maintain your non-standard parameters, since any negligence will be punished by un-predicted behavior!

In order to use the non-standard parameters, \GLOBES\ creates any set of oscillation parameters (type {\tt glb\_params})
with the additional parameters. To access the non-standard parameters, you can use the following functions as usual:
\begin{itemize}
\item
{\tt glbSetOscParams}
\item
{\tt glbGetOscParams}
\item
{\tt glbSetProjectionFlag}
\item
{\tt glbGetProjectionFlag}
\end{itemize}
In all these cases, the last parameter {\tt which} can now run from $0$ to {\tt GLB\_OSCP}$-1$. For instance,
\begin{quote}
{\tt
 glbDefineParams(true\_values,0.55,0.15,0.78,0.0,0.000082,0.0022); \\
 glbSetOscParams(true\_values,0.0,6);
}
\end{quote}
sets all of the oscillation parameters if {\tt GLB\_OSCP}=7. In addition, the function {\tt glbCopyParams}
can be used to copy all parameters inluding the non-standard ones. For projections,
do not use any of the {\tt glbChi...} functions anymore except from {\tt glbChiNP} in order to have a
predictive behavior of the extra dimenstions. Similar to the oscillation parameters, you can define
the marginalization over the extra parameter(s) by {\tt glbSetProjectionFlag}. You can find a simple
example using non-standard physics on page~\pageref{ex:nsphysics}.

\example{Non-standard process on Hamiltonian level}{
\label{ex:nsphysics}
\index{norm}{Hamiltonian}
Let us discuss an example of a very simple Hamiltonian-level non-standard flavor-changing effect which 
couples $\nu_\mu$ to $\nu_\tau$ in flavor base. This effect can be parameterized
relative to the standard matter effect acting on $\nu_e$ by an additional
parameter, which we adopt as our 7th oscillation parameter. In order
to incorporate this effect, we change a few lines in the function {\tt ComputeMixingMatrix}
in {\tt glb\_probability.cc}:
\begin{quote}
{\tt {\footnotesize
double msw = 0.1*1.5*anti\_neutrinos*(10E-4)*neutrino\_energy*0.5*matter\_density; \\
md[0][0] += msw; \\
md[1][2] += msw*nsparams[0]; md[2][1] += msw*nsparams[0];
}}
\end{quote}
Here we have assume a real contribution which has to be added in a Hermitian way (the Hamiltonian
has to be Hermitian), where the first line is the standard matter effect acting on $\nu_e$ only 
($2 \sqrt{2} G_F n_e E$ expressed in our units).

In the application software, we can now use this additional parameter. Let us assume that we
want to compute the projection onto the $\stheta$-axis when we do not know the size of the  non-standard
effect, \ie, we keep it completely free but assume it to be zero as null hypothesis:
\begin{quote}
{\tt {\footnotesize
 \#define MY\_FV 6 \\
 ... \\
 glbDefineParams(true\_values,theta12,theta13,theta23,deltacp,sdm,ldm); \\
 glbSetOscParams(true\_values,0.0,MY\_FV); \\
 glbDefineParams(input\_errors,0.1*theta12,0,0,0,0.1*sdm,0); \\
 glbSetOscParams(input\_errors,0,MY\_FV);  \\
 glbSetDensityParams(input\_errors,0.05,GLB\_ALL); \\
 glbSetOscillationParameters(); glbSetRates(); \\
 glbSetStartingValues(true\_values);glbSetInputErrors(input\_errors); \\
 \\
 glbDefineProjection(my\_proj,GLB\_FREE,GLB\_FIXED,GLB\_FREE,\\
  \hspace*{2cm} GLB\_FREE,GLB\_FREE,GLB\_FREE); \\
 glbSetProjectionFlag(my\_proj,GLB\_FREE,MY\_FV); \\
 glbSetDensityProjectionFlag(my\_proj,GLB\_FREE,GLB\_ALL); \\
 glbSetProjection(my\_proj);
 \\
 glbCopyParams(true\_values,fit\_values); \\
 glbSetOscParams(fit\_values,0.1,GLB\_THETA\_13); \\
 res=glbChiNP(fit\_values,minimum,GLB\_ALL); 
}}
\end{quote}
This code fragment illustrates the implementation of a {\tt glbChiTheta} function
including the non-standard effect, which is being marginalized over.
Alternatively, one could define a {\tt glbChiTheta} fixing the non-standard
parameter, or include external constrainys to limit its size. Therefore, 
there is no ``standard'' treatment of new physics effects incorporated
in \GLOBES .
}
